---
title: "resi_heights_processing"
author: "abbey yatsko"
date: "2023-02-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library prep
```{r include = FALSE}
library(tidyverse)
library(future.apply)
library(dplyr)
library(plyr)
library(tidyr)
library(stringr)
library(ggpubr)
library(ggplot2)
```

load functions 
```{r}
# function to read residrill data in from .rgp files 
read_resi <- function(x) {
  #Extract drill data
  drill <- x[grep("drill",x = x)]
  drill <-gsub(pattern ='"|:|\\s|\\[|\\]|[[:alpha:]]|\\,$',replacement = "",x = drill,perl =TRUE)
  drill <- as.double(unlist(strsplit(drill,split = ",")))
  
  #extract feed data
  feed <- x[grep("[[:punct:]]{1}feed[[:punct:]]{1}",x = x)]
  feed <-gsub(pattern ='"|:|\\s|\\[|\\]|[[:alpha:]]|\\,$',replacement = "",x = feed,perl =TRUE)
  feed <- as.double(unlist(strsplit(feed,split = ",")))
  
  #get observation id
  id <- x[grep("idNumber",x = x)]
  id <-gsub(pattern = "idNumber|[[:punct:]]|\\s",replacement ="" ,x = id)
  
  #resi number
  number <- x[grep("[[:punct:]]{1}number[[:punct:]]{1}",x = x)]
  number <-gsub(pattern = "[[:alpha:]]|[[:punct:]]|\\s",replacement ="" ,x = number)
  
  #depth 
  depth_mm <- x[grep("depthMsmt",x = x)]
  depth_mm <-gsub(pattern = "[[:alpha:]]|[[:punct:]]|\\s",replacement ="" ,x = depth_mm)
  depth_mm <- as.double(depth_mm)
  depth_mm <- 1:depth_mm
  out <- data.frame(id, number, drill, feed, depth_mm)
  
  return(out)
}

# function to find inflection points and generate knee_list
knee_f <- function(x){
  xs <- x$obj$K
  
  ys <- x$obj$U
  dat <- data.frame(xs, ys)
  dat <- dat[is.finite(rowSums(dat)),] ##get rid of inf values in r_list$obk
  #print(dat)
  
  knee <- uik(x = dat$xs, y= dat$ys)
  return(knee)
}

# function for dynamic programming
dynProg.mean <- function(y, Kmax, Lmin) {
  Nr  <- Kmax - 1
  n <- length(y)
  V <- matrix(Inf, nrow = n, ncol = n)
  for (j1 in (1:(n-Lmin+1))){
    for (j2 in ((j1+Lmin-1):n)) {
      yj <- y[j1:j2]
      nj <- j2-j1+1
      V[j1,j2] <- sum(yj^2) - (sum(yj)^2)/nj
    }
  }
  
  U <- vector(length=Kmax)
  U[1] <- V[1,n]
  D <- V[,n] 
  Pos <- matrix(nrow = n, ncol = Nr) 
  Pos[n,] <- rep(n,Nr)    
  tau.mat <- matrix(nrow = Nr,ncol = Nr) 
  for (k in 1:Nr){
    for (j in 1:(n-1)){
      dist <- V[j,j:(n-1)] + D[(j+1):n]
      D[j] <- min(dist)
      Pos[j,1] <- which.min(dist) + j
      if (k > 1) { Pos[j,2:k] <- Pos[Pos[j,1],1:(k-1)] }
    }
    U[k+1] <- D[1]
    tau.mat[k,1:k] <- Pos[1,1:k]-1
  }
  out <- list(Test=tau.mat, obj=data.frame(K=(1:Kmax),U=U))
  return(out)
}

# function to output sound wood segment classification 
get_sound <- function(df, r, Kopt, max_lenght=NULL) {
  y <- df$feed
  n <- length(y)
  Topt <- c(0,r$Test[(Kopt-1),1:(Kopt-1)],n)
  Tr <- c(0,Topt[2:Kopt],n) 
  dm <- data.frame()
  for (k in (1:Kopt)) {
    m <- median(y[(Topt[k]+1):Topt[k+1]],na.rm = TRUE)
    sd <- mad(y[(Topt[k]+1):Topt[k+1]],na.rm = TRUE)
    dm <- rbind(dm, c(Tr[k],Tr[k+1],m,sd))
  }
  names(dm) <- c("x1","x2","mu", "sd")
  return(dm)
}

# function to calculate and remove bark from each trace 
bark_f <- function(df) {
  df$bark <- floor((unique(df$bark)*10)) # convert bark in mm to 1/10 of mm (like residrill data)
  df$bark_far <-  df$bark+100 # added another cm to the bark on the far side of the tree as often resi would keep going before retracting
  near <- df$bark[1] # mm on which first bark ends (entry point)
  far <- df$bark_far[1] # mm on which second bark starts (exit point)
  length <- length(df$bark)

    if (df$abs_diff[1] <= 5) {
    # for complete trace capturing entirety of tree dbh
    out <- df[near:(length-far),]
    
  } else{
    # detect anomalies if resi did not come out the other end (incomplete trace)
    out <- df[near:length,]
  }
  out$depth_mm_nb <- 1:nrow(out)
  out
}

# function to split tree in half 
tree2half <- function(x) {
  # calculate radial distance
  x$radial <- x$depth_mm_nb-unique((x$drill_dbh_cm*100)/2)
  x$half <- ifelse(x$radial>0, yes = "upper",no = "lower")
  # convert all values to positive to analyse
  x$radial <- abs(x$radial)
  # rearrange tables into radial distance (i.e. starting from center for both lower and upper halves)
  x <- arrange(x,radial)
}

# function to fix traces < 90% coverage: splits tree in half, selects first half, and mirrors first half to recreate second half
fix_tree_half <- function(x) {
  # calculate radial distance
  x$radial <- x$depth_mm_nb-unique((x$drill_dbh_cm*100)/2)
  x$half <- ifelse(x$radial>0, yes = "upper",no = "lower")
  
  # remove 'upper' rows as this is an incomplete half 
  x <- x[!x$half == "upper",]
  
  # duplicate lower half to mirror as replacement for incomplete upper half 
  y <- x
  y$half[y$half == "lower"] <- "upper"
  
  # flip upper df 
  y <- y[order(-y$radial),]
  
  # remove radial = 0 (there can only be one and lower gets it)
  y <- y[!y$radial == 0, ]
  
  # fix numbering of resi_depth
  recount <- max(x$depth_mm_nb)+1
  end <- recount + nrow(y) - 1
  y$depth_mm_nb <- recount:end
  
  # rbind true first half (lower) and synthesized second half (upper)
  df <- rbind(x, y)
  
  # convert all values to positive to analyse
  df$radial <- as.numeric(df$radial)
  df$radial <- abs(df$radial)
  
  return(df)
}

# function to get segments for sound wood classification 
get_segment <- function(resi_dat, r, Kopt, alpha, sw) {
  y <- resi_dat$feed
  n <- length(y)
  Topt <- c(0,r$Test[(Kopt-1),1:(Kopt-1)],n)
  Tr <- c(0,Topt[2:Kopt],n)
  dm <- data.frame()
  for (k in (1:Kopt)) {
    m <- median(y[(Topt[k]+1):Topt[k+1]],na.rm = TRUE)
    sd <- mad(y[(Topt[k]+1):Topt[k+1]],na.rm = TRUE)
    dm <- rbind(dm, c(Tr[k],Tr[k+1],m,sd))
  }
  names(dm) <- c("x1","x2","mu", "sd")
  
  mean_feed <- sw$mu
  sigma <- sw$sd
  dm$pc_length <- (dm$x2-dm$x1)/last(dm$x2)
  dm$z<- (dm$mu-mean_feed)/(sigma)
  dm$p_val<- pnorm(dm$z,lower.tail = TRUE)
  dm$sig <- ifelse(dm$p_val<= alpha, "*","")
  dm$x1 <- n - dm$x1 # fix the problem of intervals not matching depth_mm_nb
  dm$x2 <- n - dm$x2 # fix the problem of intervals not matching depth_mm_nb
  
  #bark beggining is not significant
  #dm$sig[1] <- "" take this out because bark has already been cut off
  return(dm)
}

# function to get segments and classify damage on tree halves
half_tree_segment <- function(resi_dat, r, Kopt, alpha, feed_tree, sw) {
  y <- resi_dat$feed
  y_all <- feed_tree$feed
  dbh_last <- unique(feed_tree$drill_dbh_cm)*100
  dbh <- seq_along(1:dbh_last)
  n <- length(y)
  Topt <- c(0,r$Test[(Kopt-1),1:(Kopt-1)],n)
  Tr <- c(0,Topt[2:Kopt],n)
  dm <- data.frame()
  for (k in (1:Kopt)) {
    m <- median(y[(Topt[k]+1):Topt[k+1]],na.rm = TRUE)
    sd <- mad(y[(Topt[k]+1):Topt[k+1]],na.rm = TRUE)
    dm <- rbind(dm, c(Tr[k],Tr[k+1],m,sd))
  }
  names(dm) <- c("x1","x2","mu", "sd")
  
  # area of semi circle
  dm$area_sc <- (pi/2)*((dm$x2/2)^2)
  # area of circle
  
  # area of the doughnut
  dm$doughtnut_area <- c(dm$area_sc[1],sapply(seq_len(Kopt-1), function(x)
    dm$area_sc[x+1]-dm$area_sc[x]))  
  # proportion area compare to whole tree
  dm$doughnut_pc_area <-(dm$doughtnut_area)/(pi*(length(y)/2)^2)
  # mean_feed <- median(feed_all)
  mean_feed <- sw$mu
  # sigma <- mad(feed_all)
  sigma <- sw$sd
  dm$pc_length <- ((dm$x2-dm$x1)/length(y))*(length(y)/length(dbh))
  dm$z<- (dm$mu-mean_feed)/(sigma)
  dm$p_val<- pnorm(dm$z,lower.tail = TRUE)
  dm$sig <- ifelse(dm$p_val<= alpha, "*","")
  
   ifelse(resi_dat$half == "upper", dm$sig[length(dm$sig)] <- "", dm$sig[1] <- "") # sapwood correction, best solution
   
  return(dm)
}

# function to fix traces to correct feed values = 0
fill_zeros <- function(x){
  # define length of 50% tree diameter 
  half_length_mm <- as.numeric((x$drill_dbh_cm / 2 )*100)[1]
  
  # find current length 
  length <- as.numeric(nrow(x))
  
  # find how many rows have to be added 
  add <- half_length_mm - length
  add <- c(1:add)
  added <- data.frame(matrix(NA, ncol=ncol(x), nrow=length(add)))
  colnames(added) <- colnames(x)
  
  # rbind added to input 
  df <- rbind(x, added)
                  
  # if x$ feed is empty, then fill with 0 
  df$feed[is.na(df$feed)] <- 0 
  
  # adjust depth of trace
  df$depth_mm <- c(1:half_length_mm)
    
  # then fill out the rest of the columns til the end based on the last record 
  out <- df %>% fill(names(df))
  
  return(out)
}
```

1. preprocessing of residrill files 
  * read .rgp files in using "read_resi" function 
  * format and merge with metadata 
  * seperate out full traces from incomplete traces (not all of the traces passed through 100% of the tree)
  * remove bark from each trace
  * generate list of all traces and working df with traces from list combined 
```{r}
# read in residrill files (.rgp format)
files <- list.files("/Users/abbeyyatsko/Desktop/repos/residrill_MS/data/residrill_traces_rgp" ,full.names = TRUE, pattern = "rgp",recursive = TRUE)

# generate large list, each element is a residrill file
rgp_in <- sapply(files, readLines, USE.NAMES= TRUE, simplify = FALSE) 

# make residrill files into data frames using the read_resi function, further formatting
resi_dat <- sapply(seq_along(rgp_in), function(x) read_resi(rgp_in[[x]]), USE.NAMES= TRUE, simplify = FALSE)
names(resi_dat) <- names(rgp_in)
resi_dat_l <- bind_rows(resi_dat)
resi_dat_l$number<- as.numeric(resi_dat_l$number)

# read in drill metadata
dat <- read.csv("/Users/abbeyyatsko/Desktop/repos/residrill_MS/data/residrill_traces.csv")

# create column recording depth of residrill path 
depth <- sapply(resi_dat, function(x) tail(x$depth_mm,1)/100,USE.NAMES = TRUE,simplify = FALSE)  
names(depth) <- names(rgp_in)

# calculate depth that residrill went to for each trace
depth_v <- t(do.call(cbind.data.frame, depth))
id <- rownames(depth_v)
depth_df <- tibble(id, depth_v)

# extract trace ID from filename
depth_df$id <- qdapRegex::ex_between(depth_df$id, "renamed/", ".rgp")

# make sure drill IDs are factor variables
dat$drill_ID <- as.factor(dat$drill_ID)
depth_df <- as.data.frame(lapply(depth_df, unlist))
depth_df$id <- as.factor(depth_df$id)

# join resi path depth with metadata
d2 <-left_join(dat,depth_df, by = c("drill_ID"="id"))

# calculate difference between resi reading depth and dbh 
d2$abs_diff <- (d2$drill_dbh_cm-d2$depth_v)

# calculate amount of bark based on Rosell 
# https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.13889
d2$bark <-d2$drill_dbh_cm^0.700

# d2 and resi_dat_l have IDs that are slightly different (PNW-01 / PNW01): from d2 remove '-' from drill_ID
d2$drill_ID <- gsub('-','',d2$drill_ID)

# separate out complete/complete traces based on percentage of trace measured based on field dbh 
d2$perc_thru <- ((d2$depth_v / (d2$drill_dbh_cm )*100))
# table(d2$perc_thru > 90)
d2$full_trace <- ifelse(d2$perc_thru > 90, "Y", "N")

# remove 3 traces that have less than 50% coverage of full trace (these will have to be special cases)
d2$full_trace <- ifelse(d2$perc_thru < 50, yes = "R", no = d2$full_trace)

# classify complete and incomplete working dfs 
complete <- d2[d2$full_trace == "Y",]
complete_join <- left_join(complete, resi_dat_l, by = c("drill_ID"="id"))
complete_join$drill_ID <- as.factor(complete_join$drill_ID) 
```

COMPLETE TRACES
```{r}
# generate full list of complete traces
complete_join %>%
  group_by(drill_ID, tree_ID, species) %>%
  mutate(max_x = max(depth_v), max_y=max(feed),
         sampled_stem = (depth_v*100)/drill_dbh_cm) %>% 
  ungroup() %>%
  group_split(drill_ID, species, drill_dbh_cm) -> resi_dat_c

# save(resi_dat_c, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/resi_dat_c.RData")

# calculate bark 
resi_dat_c_nb <- lapply(seq_along(resi_dat_c), function(x) bark_f(resi_dat_c[[x]]))
# save(resi_dat_c_nb, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/resi_dat_c_nb.RData")
```

INCOMPLETE/REMOVED TRACES
```{r}
# traces that are 50% < X < 90%
incomplete <- d2[d2$full_trace == "N",]
incomplete_join <- left_join(incomplete, resi_dat_l, by = c("drill_ID"="id"))
incomplete_join$drill_ID <- as.factor(incomplete_join$drill_ID) 

# generate full list of incomplete traces
incomplete_join %>%
  group_by(drill_ID, tree_ID, species) %>%
  mutate(max_x = max(depth_v), max_y=max(feed),
         sampled_stem = (depth_v*100)/drill_dbh_cm) %>% 
  ungroup() %>%
  group_split(drill_ID, species, drill_dbh_cm) -> resi_dat_inc
# save(resi_dat_inc, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/resi_dat_inc.RData")

# calculate bark 
resi_dat_inc_nb <- lapply(seq_along(resi_dat_inc), function(x) bark_f(resi_dat_inc[[x]]))
# save(resi_dat_inc_nb, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/resi_dat_inc_nb.RData")

# remove traces with <50% coverage from processing pipeline: deal with these seperately on a case-by-case
removed <- d2[d2$full_trace == "R",]
removed_join <- left_join(removed, resi_dat_l, by = c("drill_ID"="id"))
removed_join$drill_ID <- as.factor(removed_join$drill_ID) 

removed_join %>%
  group_by(drill_ID, tree_ID, species) %>%
  mutate(max_x = max(depth_v), max_y=max(feed),
         sampled_stem = (depth_v*100)/drill_dbh_cm) %>% 
  ungroup() %>%
  group_split(drill_ID, species, drill_dbh_cm) -> resi_dat_rem

# save(resi_dat_rem, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/resi_dat_rem.RData")

# see whats going on with these highly damaged ones 
ggplot(resi_dat_rem[[2]], aes(x= depth_mm))+
  ylab("Resistance")+
  geom_area(aes(y = feed), fill = "black", alpha = 0.5)+
  geom_area(aes(y = drill), fill ="green", alpha = 0.5)
# View(resi_dat_rem[[2]])

# VERDICT: looks like 1 and 3 were drill overloads, 13, 30% coverage. need to drop from dataset
# 2 could be filled in with zeros to fix (case: too hollow), plus its practically at 50% (45% thru)

# remove bark 
resi_dat_rem_nb <- lapply(seq_along(resi_dat_rem), function(x) bark_f(resi_dat_rem[[x]]))
# save(resi_dat_rem_nb, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/resi_dat_rem_nb.RData")
```

1.1. address incomplete traces 
  * using redi_dat_inc_nb
  * firest split tree in half
  * select first half only 
  * then mirror first half to recreate second half and generate 'complete' trace
  * fill zeros for samples
```{r}
# visualize incomplete
# pdf(file = "/Users/abbeyyatsko/Downloads/residrill_incomplete.pdf", onefile = TRUE)
# # graphing portion
# lapply(seq_along(resi_dat_inc), function(x){ggplot(resi_dat_inc[[x]], aes(x= depth_mm))+
#     ylab("Resistance")+
#     geom_area(aes(y = feed), fill = "black", alpha = 0.5)+
#     geom_area(aes(y = drill), fill ="green", alpha = 0.5)})
# dev.off()

# fix removed traces: only [[2]] fill with zeros
removed_fixed <- lapply(seq_along(resi_dat_rem_nb), function(x) fill_zeros(resi_dat_rem_nb[[x]]))

# viz check
ggplot(removed_fixed[[3]], aes(x= depth_mm_nb))+
  ylab("Resistance")+
  geom_area(aes(y = feed), fill = "black", alpha = 0.5)+
  geom_area(aes(y = drill), fill ="green", alpha = 0.5)

# remove elements 1,3 as they have to be thrown out (drill overloads)
removed_fixed <- removed_fixed[-c(1, 3)]

# merge this in with the incomplete traces to continute processing 
incomplete_to_complete <- append(removed_fixed, incomplete_fixed)

# halve and mirror the incomplete trees 
incomplete_fixed <- lapply(seq_along(incomplete_to_complete), function(x) fix_tree_half(incomplete_to_complete[[x]]))

# bring complete and fixed incomplete back into the same list
resi_dat_all_nb <- append(resi_dat_c_nb, incomplete_fixed)  
# save(resi_dat_all_nb, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/resi_dat_all_nb.RData")
```

2. calculate sound wood thresholds for entire brooklyn df
  * reorder traces from high to low feed 
  * run through dynamic programming 
  * split into 2 segments
  * define sound wood segments, mu, and sd
  * save sound wood threshold mean and sd
```{r}
# only working with complete traces to start
load("resi_dat_all_nb.RData")

# reorder all traces from high to low 
list <- lapply(seq_along(resi_dat_all_nb), function(x)
  resi_dat_all_nb[[x]][order(resi_dat_all_nb[[x]]$feed, decreasing = TRUE),]) 

# dynamic programming 
Kmax <- 5 # maximum number of segments that trace can be divided into (we ultimately only want 2 for sw classification)
Lmin <- 10  # minimum length of a segment (start at 10 originally, run at higher numbers to speed things up) 
plan(multisession) # when you want to run things in parallel (future_sapply)

r_list1 <- lapply(seq_along(list), function(x) dynProg.mean(list[[x]]$feed, Kmax, Lmin))

# set knee list (static, not optimized)
knee_list <- 2

# get sound wood segments 
sws_c_nb <- lapply(seq_along(list), function(x) get_sound(list[[x]], r_list1[[x]], knee_list))

# get mean and sd of sound wood to use as threshold
sw_c_nb <- lapply(seq_along(sws_c_nb), function(x) (sws_c_nb[[x]][1,3:4]))

# save sound wood threshold mean and sd
# save(sw_c_nb, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/sw_c_nb.RData")
```

3. split traces in half
  * split each trace into lower/upper halves using "tree2half"
  * to account for radial distances, read traces from inward to outward points (start at tree center, go out towards bark)
```{r}
# load sound wood thresholds for full traces only
load("sw_c_nb.RData")

# load residrill datasetfor all traces
load("resi_dat_all_nb.RData")

# split the tree data into two DF in a list (for lower and upper), calculate radial distance (moving from tree center outwards)
resi_dat_c_area <- lapply(resi_dat_all_nb, function(x){out <- tree2half(x); return(out)}) 

# split lists into lower and upper sub list, order each lower/upper section from center of tree to perimeter
tree_half_all <- sapply(resi_dat_c_area, function(x) out <- split(x = x, f = paste0(x$half,"-",x$drill_ID)), 
                        simplify = FALSE,USE.NAMES = TRUE)

# keep the total feed as part of each half as it is needed for analysing both segments
full_feed <- rep(resi_dat_c_area,each = 2)

# name the vectors of full feed, split by unique drill ID and then upper/lower segment
names(full_feed) <- paste0(list("lower-","upper-"), sapply(full_feed, function(x) unique(x$drill_ID)))

# replicate sound wood df (2 for each residrill trace)
sound_wood <- rep(sw_c_nb,each = 2)

# name the vectors of sound wood
names(sound_wood) <- names(full_feed)

# each tree lower/upper now in its own list for processing (currently it is a list of lists)
tree_half_all <- unlist(tree_half_all, recursive = FALSE, use.names = TRUE)

# have sw list matched to each tree as its separated out by lower/upper 
sound_wood <- sound_wood[names(sound_wood) %in% names(tree_half_all)]

# filter full feed to match 
full_feed <- full_feed[names(full_feed) %in% names(tree_half_all)]
```

4. classify damage on tree half segments
  * run dynamic programming seperately on lower/upper segments 
  * define optimal inflection points 
  * classify damage on optimal segments
```{r}
# run dynamic programming 
# set parameters
Kmax <- 50 # maximum number of segments 
Lmin <- 10  # minimum length of a segment (work with 10) 
plan(multisession)

r_list <- lapply(seq_along(tree_half_all), function(x) dynProg.mean(tree_half_all[[x]]$feed, Kmax, Lmin))

# find optimal breakpoints 
knee_list <- lapply(r_list, function(x) knee_f(x)) 

# figure out output of half_tree_segment 
area_mat <- sapply(seq_along(tree_half_all), function(x) half_tree_segment(tree_half_all[[x]], r_list[[x]], knee_list[[x]], alpha = 0.05, full_feed[[x]], sound_wood[[x]]), simplify = FALSE, USE.NAMES = TRUE )

# keep names of traces
names(area_mat) <- names(tree_half_all)
```

5. DATA OUT: quantify damage
```{r}
damage_area <- sapply(area_mat, function(x) sum(x$doughnut_pc_area[which(x$sig=="*")]),simplify = FALSE,USE.NAMES = TRUE)
damage_area_df <- do.call(rbind, damage_area)

damage_length <- sapply(area_mat, function(x) sum(x$pc_length[which(x$sig=="*")]),simplify = FALSE,USE.NAMES = TRUE)
damage_length_df <- do.call(rbind, damage_length)

dat_compare <- data.frame(cbind(area = damage_area_df, linear = damage_length_df))
dat_compare <- cbind(resi_half = rownames(dat_compare), dat_compare)
colnames(dat_compare) <- c("resi_half", "area", "linear")
rownames(dat_compare) <- NULL

# export full trace first estimates 
# write.csv(dat_compare, file="/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/data_out/damage_est_c.csv")
```
