---
title: "draft_models"
author: "abbey yatsko"
date: "2023-04-18"
output: html_document
---

libraries
```{r}
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(lme4)
library(easystats)
library(multcompView)
library(ggeffects)
library(nlme)
library(stats)
library(car)
library(ggpmisc)
library(glmmTMB)
library(patchwork)
library(BIOMASS)
```

data load 
```{r}
load("/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/tree_level_damage_myrt.RData")
load("/Users/abbeyyatsko/Desktop/repos/residrill_MS/R/Rdata/myrt.RData")

agb <- read.csv("/Users/abbeyyatsko/Desktop/repos/residrill_MS/data/site_agb.csv")
pnw_stems <- read.csv("/Users/abbeyyatsko/Desktop/repos/residrill_MS/data/pnw_biomass.csv")
```

Figure 2 
```{r}
# a. relationship between proportion damage and tree species
myrt$species <- fct_relevel(myrt$species, "Corymbia clarksoniana", "Eucalyptus cullenii", "Melaleuca stenostachya")

m1 <- glmmTMB(linear_mean ~ drill_height_cm + dbh_cm.x + species + (1|tree_site_ID), myrt, family=list(family="beta",link="logit"))
summary(m1)
check_model(m1)
Anova(m1)

prop1 <- ggpredict(m1, terms = c("species")) 
a <- plot(prop1)+
   labs(x = "Species", y = "Proportion damaged", title = "Predicted proportion damage")

# b. relationship between area damage and tree species
# define area of slice 1cm thick at each measurement point 
myrt$V_xsection_cm3 <- ((myrt$drill_dbh_cm/2)^2)*(pi)*(1)
myrt$V_xsection_cm3_damaged <- myrt$V_xsection_cm3 * myrt$linear_mean

m2 <- lmer(log(V_xsection_cm3_damaged) ~ drill_height_cm + species + dbh_cm.x + (1|tree_site_ID), REML = F, data = myrt)
summary(m2)
Anova(m2)
check_model(m2)

area <- ggpredict(mixed.M1, terms = c("species")) 
b <- plot(area)+
   labs(x = "Species", y = "Absolute damaged area (cm2)", title = "Predicted absoute area damage") 

# c. relationship between proportion damage and height in tree by size 
c <- ggplot(myrt, aes(x = drill_height_cm, y = linear_mean)) + 
  geom_jitter() +
  facet_wrap(~species) +
  theme_classic() + 
  xlab("Drill height (cm)") + 
  ylab("Proportion damaged")

# d. relationship between area damage and height in tree by size 
v <- c(10, 25, 34)
area <- ggpredict(m2, terms = c("drill_height_cm [all]", "dbh_cm.x [v]"), back.transform = TRUE)  #%>% plot(add.data = TRUE)
low <- "#364B9A"
mid <- '#F99858'
high <- '#A50026'

d <- ggplot(myrt, aes(x = drill_height_cm, y = V_xsection_cm3_damaged, color = dbh_cm.x)) + 
  geom_jitter() +
  facet_wrap(~species) +
  scale_colour_gradient2(low=low, mid=mid, midpoint = 30 , high=high, name ='DBH') +
  geom_line(aes(x=x, y=predicted), data=filter(area, group=='10'), 
            inherit.aes=FALSE, colour= low) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(area, group=='10'), 
              inherit.aes=FALSE,  alpha=0.2, fill = low) +
  geom_line(aes(x=x, y=predicted), data=filter(area, group =='25'), 
            inherit.aes=FALSE, colour= mid) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(area, group=='25'), 
              inherit.aes=FALSE,  alpha=0.2, fill = mid)+
  geom_line(aes(x=x, y=predicted), data=filter(area, group =='34'), 
            inherit.aes=FALSE, colour= high) + 
  geom_ribbon(aes(x=x, ymin=conf.low, ymax=conf.high), data=filter(area, group=='34'), 
              inherit.aes=FALSE,  alpha=0.2, fill = high)+
  theme_classic() + 
  xlab("Drill height (cm)") + 
  ylab("Absolute damaged area (cm2)")


a+b+c+d + patchwork::plot_annotation(tag_levels = "a")
```

Figure 2 REVISED (a-b is about species, c-d is about traits)
```{r}
# a. relationship between proportion damage and tree species
m.1 <- glmmTMB(linear_mean ~ drill_height_cm  + species + (1|tree_site_ID), myrt, family=list(family="beta",link="logit"))
summary(m.1)
check_model(m.1)
Anova(m.1)

prop1 <- ggpredict(m.1, terms = c("species")) 
a <- plot(prop1)+
   labs(x = "Species", y = "Proportion damaged", title = "Predicted proportion damage")

# b. relationship between area damage and tree species
m.2 <- lmer(log(V_xsection_cm3_damaged) ~ drill_height_cm + species + (1|tree_site_ID), REML = F, data = myrt)
summary(m.2)
Anova(m.2)
check_model(m.2)

area <- ggpredict(m.2, terms = c("drill_height_cm", "species")) 
b <- plot(area)+
   labs(x = "Species", y = "Absolute damaged area (cm2)", title = "Predicted absoute area damage") 

# c. relationship between proportion damage and height in tree by size 
m.3 <- glmmTMB(linear_mean ~ drill_height_cm  + dbh_cm.x + wd + (1|tree_site_ID), myrt, family=list(family="beta",link="logit"))
summary(m.3)
check_model(m.3)
Anova(m.3)

prop1 <- ggpredict(m.3, terms = c("wd", "dbh_cm.x")) 
c <- plot(prop1)+
   labs(x = "Wood density", y = "Proportion damaged", title = "Predicted proportion damage")

# d. relationship between area damage and height in tree by size 
m.4 <- lmer(log(V_xsection_cm3_damaged) ~ drill_height_cm + dbh_cm.x + wd + (1|tree_site_ID), REML = F, data = myrt)
summary(m.4)
Anova(m.4)
check_model(m.4)

area <- ggpredict(m.4, terms = c("drill_height_cm", "dbh_cm.x")) 
d <- plot(area)+
   labs(x = "", y = "Absolute damaged area (cm2)", title = "Predicted absoute area damage") 

a+b+c+d + patchwork::plot_annotation(tag_levels = "a")
```

Figure 3 
Tapering - volume damaged using frustrum 
```{r}
# generate list of all trees 
ids <- unique(myrt$tree_site_ID)

# function that calculates undamaged frustrum volume 
frus_ud <- function(x){
  ID <- unique(x$tree_site_ID)
  d1 <- min(x$drill_height_cm) 
  d4 <- max(x$drill_height_cm)
  L <- d4 - d1
  A1 <- x$V_xsection_cm3[x$drill_height_cm == d1]
  A4 <- x$V_xsection_cm3[x$drill_height_cm == d4]
  V <- (L/3) * (A1 + sqrt(A1 * A4) + A4)
  
  output <- data.frame(tree_site_ID = ID, frus_undamage_cm3 = V)
  return(output)
}

# function that calculates damaged frustrum volume 
frus_d <- function(x){
  ID <- unique(x$tree_site_ID)
  d1 <- min(x$drill_height_cm) 
  d4 <- max(x$drill_height_cm)
  L <- d4 - d1
  A1 <- x$V_xsection_cm3_damaged[x$drill_height_cm == d1]
  A4 <- x$V_xsection_cm3_damaged[x$drill_height_cm == d4]
  V <- (L/3) * (A1 + sqrt(A1 * A4) + A4)
  
  output <- data.frame(tree_site_ID = ID, frus_damage_cm3 = V)
  return(output)
}

# calculate volume of undamaged frustrum for each tree 
frus_v <- lapply(ids, function(id) frus_ud(myrt[myrt$tree_site_ID == id, ]))
frus_v_out <- do.call(rbind, frus_v)  

# calculate volume of damaged frustrum for each tree 
frus_v_d <- lapply(ids, function(id) frus_d(myrt[myrt$tree_site_ID == id, ]))
frus_v_d_out <- do.call(rbind, frus_v_d)  

# merge frustrum volume estimates for damaged and undamaged 
frustrum <- merge(frus_v_out, frus_v_d_out, by = "tree_site_ID")

# calculate biomass using wood density
info <- tree_level_damage_myrt[, c("tree_site_ID", "species", "wd", "dbh_cm", "site")]
frustrum <- merge(info, frustrum, by = "tree_site_ID")

# MULTIPOINT damage estimate
frustrum$undamaged_kg <- (frustrum$wd * frustrum$frus_undamage_cm3)/1000
frustrum$mp_damaged_kg <- (frustrum$wd * frustrum$frus_damage_cm3)/1000

# SINGLEPOINT damage estimate (from residrill closest to DBH_)
dbh_resi <- myrt[myrt$drill_height_cm > 113 & myrt$drill_height_cm < 136, ]
missing_elements <- ids[!ids %in% dbh_resi$tree_site_ID]

# troubleshoot missing dbh residrill measurements (STCK-1 only 40cm, STCK-10 needs height = 109, STCK-18 repeated 2x)
stck10 <- myrt %>% filter(tree_site_ID == "STCK-10" & drill_height_cm == 109)
dbh_resi <- rbind(dbh_resi, stck10)
dbh_resi <- dbh_resi %>% filter(!(tree_site_ID == "STCK-18" & drill_height_cm == 114))

# merge in linear_mean of dbh measurement for each tree to frustum df
dbh_resi_merge <- dbh_resi[c("tree_site_ID", "linear_mean")]
frustrum_agb <- left_join(dbh_resi_merge, frustrum, by = "tree_site_ID")
frustrum_agb$sp_damaged_kg <- (frustrum_agb$linear_mean * frustrum_agb$undamaged_kg)

# proportion of biomass damaged from single point and multi point estimate
frustrum_agb$sp_damaged_prop <- frustrum_agb$sp_damaged_kg/frustrum_agb$undamaged_kg
frustrum_agb$mp_damaged_prop <- frustrum_agb$mp_damaged_kg/frustrum_agb$undamaged_kg
frustrum_agb_long <- pivot_longer(frustrum_agb, cols = sp_damaged_kg:mp_damaged_kg)

t <- wilcox.test(frustrum_agb$sp_damaged_kg, frustrum_agb$mp_damaged_kg, paired = TRUE)

frustrum_agb_long %>% group_by(name) %>%
  summarise(mean = mean(value), 
            sd = sd(value)) -> df 

ggplot(df, aes(x = name, y = mean, fill = name))+ 
  geom_bar(stat = "identity", position = "dodge")+
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0.07)+
  theme_classic()+
  ylab("AGB in frustrum (kg)")+
  xlab("")+
  theme(legend.position="none")
```

Figure 4
```{r}
# merge in presence/absence data for termite/fire presence
tree_damage <- tree_level_damage_myrt[c("tree_site_ID", "site", "termite_pres", "mound", "runway", "termite", "fire")]
frustrum <- left_join(frustrum, tree_damage, by = c("tree_site_ID", "site"))
mod <- lm(mp_damaged_kg ~ fire+termite_pres, data = frustrum)
summary(mod)
check_model(mod)
pred_data <- ggpredict(mod, terms = c("termite_pres","fire"))  %>% plot(add.data = TRUE)

# plotting seperately (different than the combined model)
a <- ggplot(frustrum, aes(x = termite_pres, y = mp_damaged_kg, fill = termite_pres))+ 
  geom_boxplot()+
  theme_classic()
b <- ggplot(frustrum, aes(x = fire, y = mp_damaged_kg, fill = fire))+ 
  geom_boxplot()+
  theme_classic()
ggarrange(a, b, ncol = 2)

# at species level 
mod_sp <- lm(mp_damaged_kg ~ fire + termite_pres + species, data = frustrum)
summary(mod_sp)
check_model(mod_sp)
ggpredict(mod_sp, terms = c("termite_pres", "fire", "species"))  %>% plot(add.data = TRUE)

# at traits level 
mod_tr <- lm(mp_damaged_kg ~ fire + termite_pres + wd + dbh_cm, data = frustrum)
summary(mod_tr)
check_model(mod_tr)
ggpredict(mod_sp, terms = c("termite_pres", "fire", "dbh_cm"))  %>% plot(add.data = TRUE)

mod_t <- lm(mp_damaged_kg ~ termite_pres*wd + termite_pres*dbh_cm, data = frustrum)
summary(mod_t)

###DRAFT#### ggpredict to boxplot 
# Extract observed data
obs_data <- data.frame(termite_pres = frustrum$termite_pres,
                       fire = frustrum$fire,
                       response = frustrum$mp_damaged_kg) 
ggplot() +
# Plot predicted values as boxplot
  geom_boxplot(data = pred_data, aes(x = x, y = predicted, color = group)) +
  # Add observed data as points
  geom_point(data = obs_data, aes(x = termite_pres, y = response, color = factor(fire)), position = position_jitterdodge()) +
  # Adjust appearance (customize as needed)
  labs(x = "Termite Presence", y = "Response", color = "Fire") +
  theme_minimal() +
  theme(legend.position = "top")
```

Figure 5
```{r}
# relationships between termite/fire evidence for different size/wd  
# binomial model - define coding 
myrt$termite_num <- ifelse(myrt$termite_pres == "Y", 1, 0)
myrt$mound_num <- ifelse(myrt$mound == "Y", 1, 0)
myrt$runway_num <- ifelse(myrt$runway == "Y", 1, 0)
myrt$termitesampled_num <- ifelse(myrt$termite == "Y", 1, 0)
myrt$fire_num <- ifelse(myrt$fire == "Y", as.integer(1), as.integer(0))

# model - DBH
myrt %>%
  group_by(tree_site_ID, species, site) %>%
  distinct(dbh_cm.x, wd, fire_num, termite_num, mound_num, runway_num, termitesampled_num)-> pp

# single model with interaction between wd and size
pp$category <- fct_relevel(as.factor(pp$wd), "0.9770", "0.8453", "0.7405")
                           
test <- glm(termite_num ~ dbh_cm.x * category, data = pp, family = binomial(logit))
summary(test)
Anova(test)
r2(test)
vif(test)
check_model(test)
ggpredict(test, terms = c("dbh_cm.x","category")) %>% plot()

# two models - split out DBH and wd
term_dbh <- glm(termite_num ~ dbh_cm.x, data = pp, family = binomial(logit))
summary(term_dbh)
Anova(term_dbh)
dbh <- ggpredict(term_dbh, terms = c("dbh_cm.x"))  %>% plot(add.data = TRUE)

term_dbh_graph <- ggplot(pp, aes(x = dbh_cm.x, y =  termite_num))+
  geom_point()+
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = TRUE, color = low)+
  theme_classic()+
  xlab("DBH (cm)")+
  ylab("Termite activity")
  
term_wd <- glm(termite_num ~ wd , data = pp, family = binomial("logit"))
summary(term_wd)
Anova(term_wd)
wd <- ggpredict(term_wd, terms = c("wd")) %>% plot(add.data = TRUE)

term_wd_graph <- ggplot(pp, aes(x = wd, y =  termite_num))+
  geom_point()+
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = TRUE, color = low)+
  theme_classic()+
  xlab("Wood density (g/cm3)")+
  ylab("")

term_dbh_graph + term_wd_graph + patchwork::plot_annotation(tag_levels = "a")
```

Figure 6 (??) do bigger / denser trees have greater amounts of damage in the frustum? 
```{r}
# size model
mod_test <- lm(mp_damaged_kg ~ dbh_cm, data = frustrum)
summary(mod_test)
check_model(mod_test)
ggpredict(mod_test, terms = c("dbh_cm"))  %>% plot(add.data = TRUE)

# density model
mod_test <- lm(mp_damaged_kg ~ wd, data = frustrum)
summary(mod_test)
check_model(mod_test)
ggpredict(mod_test, terms = c("wd"))  %>% plot(add.data = TRUE)

```

Supplementary Table 1. Site description, stem percentages, and species biomass
```{r}
# station creek biomass summary
stck_agb <- agb[agb$site == "station_creek",]
stck_agb %>% group_by(species) %>%
  summarise(biomass = sum(agb), 
            count = n(), 
            agb_perc = (biomass/15.38937)*100, 
            stem_perc = (count/50)*100) -> stck_out

# pennyweight biomass calculate
pnw_stems$dbh <- pnw_stems$Circ_cm/(2*pi)
pnw_stems$wd <- NA
pnw_stems$wd <- ifelse(pnw_stems$species == "Melaleuca stenostachya", 0.7405, pnw_stems$wd)
pnw_stems$wd <- ifelse(pnw_stems$species == "Eucalyptus cullenii", 0.977, pnw_stems$wd)
pnw_stems$wd <- ifelse(pnw_stems$species == "Terminalia aridicola subsp. chillagoensis", 0.74, pnw_stems$wd)
pnw_stems$wd <- ifelse(pnw_stems$species == "Gardenia vilhelmii", 0.64, pnw_stems$wd)

pnw_stems$agb <- computeAGB(D = pnw_stems$dbh, WD = pnw_stems$wd, coord = c(144.92,-16.57))
  
# pennyweight biomass summary
pnw_stems %>% group_by(species) %>%
  summarise(biomass = sum(agb), 
            count = n(), 
            agb_perc = (biomass/2.324303)*100, 
            stem_perc = (count/49)*100) -> pnw_out
```

Supplementary Table 2. Sample sizes for residrill measurements

Supplementary Figure 1, Table E
```{r}
# relationship between dbh and wd 
# model
mod3 <- lm(dbh_cm.x ~ wd, data = pp)
summary(mod3)

ggpredict(mod3, terms = c("wd")) %>% plot(add.data = TRUE)

wd <- ggplot(pp,aes(x= wd, y = dbh_cm.x, colour= species))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_classic()
```
